Improving the Image Acquisition Rate \\ of an Atomic Force Microscope \\ Through Sub-sampling and Reconstruction

### 2019-06-30 latex/afm-cs-main.tex
Final editing tweaks.

\newcommand{\rzup}{\ensuremath{r_{z,\textrm{up}}}\xspace}
\newcommand{\rzs}{\ensuremath{r_{z,\textrm{s}}}\xspace}

@@

\author{Roger A. Braker, \textit{Student Member, IEEE}, Yufan Luo, \textit{Student Member, IEEE} , Lucy Y. Pao, \textit{Fellow, IEEE} and Sean B. Andersson, \textit{Senior Member, IEEE, and Member, ASME}
  \thanks{Braker and Pao are with the Dept. of Electrical, Computer, and
    Energy Engineering at the University of Colorado, 425 UCB,
    Boulder, CO 80309 USA. %Phone: 
    +1 (303) 492-2360.} 
    %Fax:
    %+1 (303) 492-2758.}%R. A. Braker (corresponding author
    %roger.braker@colorado.edu) is a graduate student and L.Y. Pao
    %(pao@colorado.edu) is the Palmer  Professor.}
    \thanks{Luo and Andersson are with the Division of Systems Engineering and Andersson is also with the Department of Mechanical Engineering, Boston University, Boston, MA 02215 USA.}
  \thanks{This work was supported in part by the US National Science
    Foundation (NSF Grants CMMI-1234980, CMMI-1234845, CMMI-1562031, and DBI-1352729), Agilent Technologies,
    Inc., a Palmer Endowed Chair Professorship, and the Hanse-Wissenschaftskolleg Institute for Advanced Study in Delmenhorst, Germany.} }

@@

\noindent
\begingroup \setlength{\tabcolsep}{1pt}
\begin{figure*}[ht!]
  \centering
  \begin{tabular}{cccccc}
    \textit{\small original image} & \textit{\small BPDN} & \textit{\small BPVV}
    & \textit{\small detail of original} & \textit{\small detail of BPDN} & \textit{\small detail of BPVV} \\
    \includesvg[width=0.16\textwidth]{figures/cs20ng_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp_zoom.svg}
    %
    &   \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv_zoom.svg}\\
    % 
    % ------------------------------
    % Row
    % 2
    % --------------------------------------------
    \includesvg[width=0.16\textwidth]{figures/dna_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv_zoom.svg}
  \end{tabular}
  \caption{Reconstruction comparison between BPDN and BPVV. (first
    column) Original raster-scanned image. (second column) BPDN
    reconstruction from random, $\mu$-path sampling.
    (third column) BPVV reconstruction from the same
    sub-sampled data. (remaining columns) Corresponding details from
    the red boxes indicated in the raster image.
    The top row uses 15\% sampling with 50 pixel $\mu$-paths and the
    bottom row uses 25\% sampling with 25 pixel $\mu$-paths.
    The results show that BPVV reduces the artifacts arising from the
    horizontal scans of the $\mu$-path pattern that appear in BPDN
    reconstructions.}
  \label{fig:BPTV_demonstration}
\end{figure*}
\endgroup

@@

A simulation comparison between BPDN and BPVV is shown in
Fig.~\ref{fig:BPTV_demonstration}. The first column shows two raster
scans: the top is a 512 by 512 pixel image of a CS-20NG grating (Ted Pella, USA) acquired
with an Agilent 5400 AFM and the bottom is a 256 by 256 pixel image of
DNA acquired with an Agilent 5500 AFM. For the CS-20NG grating, we
sampled 15\% of the pixels with 50 pixel $\mu$-paths and 25\%
of the DNA image with 25 pixel long $\mu$-paths. The second columns
shows the full frame reconstructions using BPDN and the third columns
shows the result using BPVV (with $\alpha = 0.75$ and $\beta = 0.1$). The fourth column zooms in on the region
indicated by the red box in column one. Similarly, the fifth and six
columns show the detail of the reconstructions. The reconstruction
results show the artifacts apparent in the BPDN reconstruction are
largely mitigated using BPVV. Thus, in the remainder of the paper, we
focus exclusively on BPVV. These reconstructions wth BPVV took about
1.5 seconds for the DNA image and about 7 seconds for the
CS-20NG (on a typical desktop computer).

@@

In order to minimize the scanning time, the tip should be moved as fast as possible when trainsitioning between $\mu$-path locations (state (1)). Since such speeds can be beyond the bandwidth of the $z-$axis controller, it is important to reduce the tip-sample interaction force to prevent tip wear and sample damage. In our initial implementation \cite{braker_hardware_2018}, 
%we insisted that the during
this was achieved by insisting that during the
tip-retraction (state (5)), the cantilever tip should fully break
contact with the surface. Here, we impose no such requirement. Rather,
we change the $Z$-setpoint during tip-retraction to only pull away far
enough that the deflection signal stays low (i.e., below the scanning
setpoint), even while we run across the surface. We choose \rzup so
that the interaction between the tip and sample becomes purely
adhesive. In general, stable imaging is not possible with the setpoint at \rzup and
complete dis-engagement occasionally occurs. This can be seen in the bottom pane of
Fig.~\ref{fig:cs_cycle} during the second $XY$-move (black portion) where the tip detaches partway through the move. When dis-engagement does occur, the $Z$-axis control loop implements an anti-windup scheme to prevent the control from saturating.

@@

In general, the $X$-axis positioning requirements are more stringent
for both CS and raster scanning. Due to the step inputs the $X$-axis
must track, we have found that compensating hysteresis and creep to be
particular helpful during $\mu$-path scanning. Thus, the $X$-axis
compensator also includes an inverse creep and hysteresis model. The
hysteresis model is a modified Prandtl-Ishlinskii type model and is
based on the work of \cite{kuhnen_modeling_2003}, while the drift
compensation models the piezo creep as a simple second-order transfer
function with real poles and zeros. In Fig.~\ref{fig:afm_bd_dinv}, the
hysteresis inversion is indicated by $\mathcal{H}_X^{-1}[\cdot]$. We
have found that the inclusion of these extra compensators limits the
overshoot of the $X$-axis during larger moves between different
$\mu$-paths (e.g., without them, the $X$-axis trajectory in
Fig.~\ref{fig:with_ff} would show noticeable overshoot). More details
on fitting the creep and hysteresis models can be found in our other
work \cite{braker_afmmpc_2019}.

@@

For raster scanning, we set $F_X=F_Y=1$ and instead take advantage of
the periodicity of the $X$-axis reference, which is a triangle wave.
We modify the triangle wave to a truncated Fourier series, where each
Fourier coefficient is scaled by the inverse of the closed-loop FRF at
the corresponding frequency. See, e.g., \cite{clayton_review_2009} for
more details on this strategy. We truncate this series such that the
highest frequency is smaller than 200 Hz, which means that we should
never excite the problematic $G_{Z,u_X}$ modes. This scheme implies
that the effective bandwidth of the $X$-axis controller is higher
while raster scanning than while CS-scanning. On the other hand, it
also implies that for faster scan rates, fewer Fourier
coefficients are used, and so the approximation to a triangle wave is
worse, leading to scans which do not cover the entire intended range.

@@

Things are clearer if we plot the RDI for each scan against total
acquisition time, shown in Fig.~\ref{fig:time_damage}.
%, and this is where CS seems to shine. 
As one would expect, as imaging time
decreases, the RDI for each scan type increases. Note that the CS curves are always better than that of the full raster scan. From
Fig.~\ref{fig:time_damage}, we see that for a fixed RDI, CS has a
faster imaging acquisition rate. The amount of improvement is best if
we need a very low RDI: for example, if we want to hold the RDI below~2, then we have to take the 1~Hz raster scan (512 seconds) while we
can achieve the same RDI via CS in about 42 seconds with the 12.5\%
sampling, which is a speed improvement of over a factor of 10. The higher of an
RDI we are willing to tolerate, the advantage of CS over raster
narrows. For example, the 8~Hz raster scan and 8~Hz, 12.5\% CS scan
have comparable RDIs, but now CS is only about 5 times faster. This is
a result of the constant overhead imposed by the engage/disengage and
$XY$-move states in CS which remains constant for different scan
rates.

@@

Fig.~\ref{fig:improve_512} compares the amount of speed improvement of
the different sub-sampling schemes over a full raster scan with 512
lines as a function of scan rate. As far as pure speed goes, the CS
scan at 12.5\% is the fastest for a rate of 1~Hz, though the 64 line
raster scan is close. At the same time, the image quality of the 1~Hz,
12.5\% CS scan is better than the 1~Hz 64 line raster scan.
This speed advantage quickly evaporates for faster scan rates.
Indeed, for rates beyond about 3~Hz, the 64 line scan is faster than the 12.5\% CS scan and
the 128 line scan is faster than the 25\% CS scan. 
In general, the artifacts of the CS reconstructions are different than the interpolated raster scans.
CS appears to preserve the edges of the holes better than sub-line sampling while sub-line sampling has less noise in the flat areas of the specimen.

>>>>

\newcommand{\rzup}{\ensuremath{r_{Z,\textrm{up}}}\xspace}
\newcommand{\rzs}{\ensuremath{r_{Z,\textrm{s}}}\xspace}

@@

\author{Roger A. Braker, Yufan Luo, \textit{Student Member, IEEE}, Lucy Y. Pao, \textit{Fellow, IEEE}\\ and Sean B. Andersson, \textit{Senior Member, IEEE, and Member, ASME}
  \thanks{Braker and Pao are with the Dept. of Electrical, Computer, and
    Energy Engineering at the University of Colorado, 425 UCB,
    Boulder, CO 80309 USA. %Phone: 
    +1 (303) 492-2360.} 
    %Fax:
    %+1 (303) 492-2758.}%R. A. Braker (corresponding author
    %roger.braker@colorado.edu) is a graduate student and L.Y. Pao
    %(pao@colorado.edu) is the Palmer  Professor.}
    \thanks{Luo and Andersson are with the Division of Systems Engineering and Andersson is also with the Department of Mechanical Engineering, Boston University, Boston, MA 02215 USA.}
  \thanks{This work was supported in part by the US National Science
    Foundation (NSF Grants CMMI-1234980, CMMI-1234845, CMMI-1562031, and DBI-1352729), Agilent Technologies,
    Inc., a Palmer Endowed Chair Professorship, and the Hanse-Wissenschaftskolleg Institute for Advanced Study in Delmenhorst, Germany.} }

@@

\noindent
\begingroup \setlength{\tabcolsep}{1pt}
\begin{figure*}[ht!]
  \centering
  \begin{tabular}{cccccc}
    \textit{\small original image} & \textit{\small BPDN} & \textit{\small BPVV}
    & \textit{\small detail of original} & \textit{\small detail of BPDN} & \textit{\small detail of BPVV} \\
    \includesvg[width=0.16\textwidth]{figures/cs20ng_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/cs20ng_bp_zoom.svg}
    %
    &   \includesvg[width=0.16\textwidth]{figures/cs20ng_bptv_zoom.svg}\\
    % 
    % ------------------------------
    % Row
    % 2
    % --------------------------------------------
    \includesvg[width=0.16\textwidth]{figures/dna_og.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_og_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bp_zoom.svg}
    %
    & \includesvg[width=0.16\textwidth]{figures/dna_bptv_zoom.svg}
  \end{tabular}
  \caption{Reconstruction comparison between BPDN and BPVV. (first
    column) Original raster-scanned image. (second column) BPDN
    reconstruction from random $\mu$-path sampling.
    (third column) BPVV reconstruction from the same
    sub-sampled data. (remaining columns) Corresponding details from
    the red boxes indicated in the raster image.
    The top row uses 15\% sampling with 50 pixel $\mu$-paths and the
    bottom row uses 25\% sampling with 25 pixel $\mu$-paths.
    The results show that BPVV reduces the artifacts arising from the
    horizontal scans of the $\mu$-path pattern that appear in BPDN
    reconstructions.}
  \label{fig:BPTV_demonstration}
\end{figure*}
\endgroup

@@

A simulation comparison between BPDN and BPVV is shown in
Fig.~\ref{fig:BPTV_demonstration}. The first column shows two raster
scans: the top is a 512 by 512 pixel image of a CS-20NG grating (Ted Pella, USA) acquired
with an Agilent 5400 AFM and the bottom is a 256 by 256 pixel image of
DNA acquired with an Agilent 5500 AFM. For the CS-20NG grating, we
sampled 15\% of the pixels with 50 pixel $\mu$-paths and 25\%
of the DNA image with 25 pixel long $\mu$-paths. The second columns
shows the full frame reconstructions using BPDN and the third columns
shows the result using BPVV (with $\alpha = 0.75$ and $\beta = 0.1$). The fourth column zooms in on the region
indicated by the red box in column one. Similarly, the fifth and six
columns show the detail of the reconstructions. The reconstruction
results show the artifacts apparent in the BPDN reconstruction are
largely mitigated using BPVV. Thus, in the remainder of the paper, we
focus exclusively on BPVV. These reconstructions with BPVV took about
1.5 seconds for the DNA image and about 7 seconds for the
CS-20NG (on a typical desktop computer).

@@

In order to minimize the scanning time, the tip should be moved as fast as possible when trainsitioning between $\mu$-path locations (state (1)). Since such speeds can be beyond the bandwidth of the $Z-$axis controller, it is important to reduce the tip-sample interaction force to prevent tip wear and sample damage. In our initial implementation \cite{braker_hardware_2018}, 
%we insisted that the during
this was achieved by insisting that during the
tip-retraction (state (5)), the cantilever tip should fully break
contact with the surface. Here, we impose no such requirement. Rather,
we change the $Z$-setpoint during tip-retraction to only pull away far
enough that the deflection signal stays low (i.e., below the scanning
setpoint), even while we run across the surface. We choose \rzup so
that the interaction between the tip and sample becomes purely
adhesive. In general, stable imaging is not possible with the setpoint at \rzup and
complete dis-engagement occasionally occurs. This can be seen in the bottom pane of
Fig.~\ref{fig:cs_cycle} during the second $XY$-move (black portion) where the tip detaches partway through the move. When dis-engagement does occur, the $Z$-axis control loop implements an anti-windup scheme to prevent the control from saturating.

@@

In general, the $X$-axis positioning requirements are more stringent
for both CS and raster scanning. Due to the step inputs the $X$-axis
must track, we have found that compensating hysteresis and creep to be
particular helpful during $\mu$-path scanning. Thus, the $X$-axis
compensator also includes an inverse creep and hysteresis model. The
hysteresis model is a modified Prandtl-Ishlinskii type model and is
based on the work of \cite{kuhnen_modeling_2003}, while the drift
compensation models the piezo creep as a simple second-order transfer
function with real poles and zeros whose inverse is part of $D_X$.
In Fig.~\ref{fig:afm_bd_dinv}, the hysteresis inversion is indicated
by $\mathcal{H}_X^{-1}[\cdot]$. We
have found that the inclusion of these extra compensators limits the
overshoot of the $X$-axis during larger moves between different
$\mu$-paths (e.g., without them, the $X$-axis trajectory in
Fig.~\ref{fig:with_ff} would show noticeable overshoot). More details
on fitting the creep and hysteresis models can be found in our other
work \cite{braker_afmmpc_2019}.

@@

For raster scanning, we set $F_X=F_Y=1$ and instead take advantage of
the periodicity of the $X$-axis reference, which is a triangle wave.
We modify the triangle wave to a truncated Fourier series, where each
Fourier coefficient is scaled by the inverse of the closed-loop FRF at
the corresponding frequency. See, e.g., \cite{clayton_review_2009} for
more details on this strategy. We truncate this series such that the
highest frequency is smaller than 200 Hz, which means that we should
never excite the problematic $G_{Z,u_X}$ modes. This scheme implies
that the effective bandwidth of the $X$-axis controller is higher
while raster scanning than while CS-scanning. On the other hand, it
also implies that for faster scan rates, fewer Fourier
coefficients are used, and so the approximation to a triangle wave is
worse, leading to scans which cover slightly less than the intended range.

@@

Things are clearer if we plot the RDI for each scan against total
acquisition time, shown in Fig.~\ref{fig:time_damage}.
%, and this is where CS seems to shine. 
As one would expect, as imaging time
decreases, the RDI for each scan type increases.
From Fig.~\ref{fig:time_damage}, we see that for a fixed RDI, CS has a
faster imaging acquisition rate. Conversely, for an acquisition time below
a given threshold, CS offers a smaller RDI than a full raster scan.
% This is not the case: e.g., compare 8Hz cs and 1 hz raster.
% Note that the CS curves are always better than that of the full raster scan.
% From Fig.~\ref{fig:time_damage}, we see that for a fixed RDI, CS has a
% faster imaging acquisition rate.
The amount of improvement is best if we need a very low RDI: for example,
if we want to hold the RDI below~2, then we have to take the 1~Hz raster scan
(512 seconds) while we can achieve the same RDI via CS in about 42 seconds with the 12.5\%
sampling, which is a speed improvement of over a factor of 10. The higher of an
RDI we are willing to tolerate, the advantage of CS over raster
narrows. For example, the 8~Hz raster scan and 8~Hz, 12.5\% CS scan
have comparable RDIs, but now CS is only about 5 times faster. This is
a result of the constant overhead imposed by the engage/disengage and
$XY$-move states in CS which remains constant for different scan
rates.

@@

Fig.~\ref{fig:improve_512} compares the amount of speed improvement of
the different sub-sampling schemes over a full raster scan with 512
lines as a function of scan rate. As far as pure speed goes, the CS
scan at 12.5\% is the fastest for a rate of 1~Hz, though the 64 line
raster scan is close. At the same time, the image quality of the 1~Hz,
12.5\% CS scan is better than the 1~Hz, 64 line raster scan.
This speed advantage quickly evaporates for faster scan rates.
Indeed, for rates beyond about 3~Hz, the 64 line scan is faster than the 12.5\% CS scan and
the 128 line scan is faster than the 25\% CS scan. 
In general, the artifacts of the CS reconstructions are different than the interpolated raster scans.
CS appears to preserve the edges of the holes better than sub-line sampling,
while sub-line sampling has less noise in the flat areas of the specimen.
